---
title: "Tidy Tuesday 2025-02-04"
author: "Zoe Carver"
format: html
editor: visual
---

```{r message=FALSE}
library(tidyverse)

simpsons_characters <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-02-04/simpsons_characters.csv')
simpsons_episodes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-02-04/simpsons_episodes.csv')
simpsons_locations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-02-04/simpsons_locations.csv')
simpsons_script_lines <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-02-04/simpsons_script_lines.csv')
```

```{r}
simpson_family <- simpsons_characters %>%
  filter(id %in% c(1, 2, 8, 9, 105)) %>% 
  mutate(char_id = id)

spoken_lines <- simpsons_script_lines %>% 
  filter(speaking_line == TRUE) %>% 
  filter(!is.na(normalized_text))

simpsons_speaking <- left_join(simpson_family, spoken_lines, by = c("char_id" = "character_id"))

simpsons_speaking <- left_join(simpsons_speaking, simpsons_episodes, by = c("episode_id" = "id"))
```

Who in the Simpson family has the most spoken lines in every season?

```{r}
simpsons_speaking %>% 
  group_by(name, season) %>% 
  summarise(lines_spoken = n()) %>% 
  complete(season = 21:26) %>% 
  mutate(lines_spoken = if_else(is.na(lines_spoken), 0, lines_spoken)) %>% 
  ggplot(aes(x = season, y = lines_spoken, color = name)) +
  geom_line(linewidth = 2) +
  scale_color_viridis_d() +
  labs(title = "Lines Spoken per Season by Simpson Family Member", color = "Character") +
  xlab("Season") +
  ylab("Lines Spoken") +
  theme_classic()

ggsave(filename = "2-04-2025_TidyTuesday.png", plot = last_plot())
```
This plot shows that, by far, Homer Simpson has the most spoken lines in the Simpson family throughout the entire show. While Marge and Homer spoke more during season 24, Bart and Lisa spoke a bit less. Maggie does not have any "spoken" lines. Bart and Lisa follow very similar trends as each other, while Marge somewhat follows Homer's path (on a lesser scale), besides season 23. 